---
title: "Reproducible Research Project 1"
author: "Shashi Ponraja"
date: "17/03/2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source ('import_data.R')
library(chron)
library (ggplot2)

```


This is an analysis of steps taken throughout the day, separated by 5 minute time intervals. It was collected between `r min(as.character(activity_data$date))` and `r max(as.character(activity_data$date))`


## Total Number of Steps Taken Per Day


```{r, echo=TRUE}
TotalStepsByDate = aggregate(steps ~ date,data = activity_data, FUN=sum)
hist(TotalStepsByDate$steps, main ='Histogram of steps per day', xlab = 'Steps per day')

```



## Median and mean steps per day

The mean number of steps taken per day is `r format(mean(TotalStepsByDate$steps),scientific=F)` and the median number of steps is `r median(TotalStepsByDate$steps)`

## Average Daily Activity Pattern

```{r, echo=TRUE}
AverageStepsByInterval = aggregate(steps ~ interval, data = activity_data, FUN = mean)
plot(AverageStepsByInterval$interval, AverageStepsByInterval$steps, type="l", xlab = '24hour time', ylab = 'steps taken', main = 'Average steps taken during the day')
```

## Interval with maximum average steps
On average, the `r AverageStepsByInterval[which.max(AverageStepsByInterval$steps),'interval',]` interval has the highest amount of steps - `r format(max(AverageStepsByInterval$steps), scientific=F)` on average


For fun, we decided to fill in all the intervals where steps weren't recorded. We used the median amount of steps for that interval to fill in missing data. This changes our dataset somewhat.


```{r, echo=TRUE}
activity_data_noNAs = activity_data
MedianStepsByInterval = aggregate(steps ~ interval, data = activity_data, FUN = median)
activity_data_noNAs[is.na(activity_data$steps),1]=MedianStepsByInterval[which(MedianStepsByInterval$'interval' == activity_data[is.na(activity_data$steps),3]),2]

noNA_TotalStepsPerDay = aggregate(steps ~ date, activity_data_noNAs, FUN = sum)
hist(noNA_TotalStepsPerDay$steps,main ='Histogram #2 of steps per day', xlab = 'Steps per day')


```

## New median / mean with modified data
Our new median is `r format(median (noNA_TotalStepsPerDay$steps),scientific=F)` and mean is `r format(mean (noNA_TotalStepsPerDay$steps), scientific=F)`. Unlike the raw unmodified dataset, our median and mean are further apart, and lower. 

## Differences in activity between weekdays and weekend


```{r, echo = TRUE}
Sys.setenv(TZ="Europe/Berlin")
activity_data_noNAs$dayType = 'weekday'
activity_data_noNAs[is.weekend(activity_data_noNAs$date),4] = 'weekend'
meanStepsDailyByDayType = aggregate(steps ~ interval + dayType, data = activity_data_noNAs,FUN=mean)

par(mfrow = c(2,1) )
#plot(meanStepsDailyByDayType[which(meanStepsDailyByDayType$dayType == 'weekday'),], )
plot(meanStepsDailyByDayType[which(meanStepsDailyByDayType$dayType == 'weekday'),]$interval, meanStepsDailyByDayType[which(meanStepsDailyByDayType$dayType == 'weekday'),]$steps, type="l", xlab = '24hour time', ylab = 'steps taken', main = 'Average steps taken during a weekday')
plot(meanStepsDailyByDayType[which(meanStepsDailyByDayType$dayType == 'weekend'),]$interval, meanStepsDailyByDayType[which(meanStepsDailyByDayType$dayType == 'weekend'),]$steps, type="l", xlab = '24hour time', ylab = 'steps taken', main = 'Average steps taken during a weekend')
```



So in these recordings, people on weekends start moving later but on average seem to move more.

